\section{The low Mach number flow equations}

TODO: This is just a place holder for a more generic description of the model, which will have to 
discuss the CHEMKIN assumptions, EGLib, etc.

\newcommand{\etal}{{\it et al.}}

\lmc\ solves the reacting navier-Stokes flow equations in the low Mach number regime.
The methodology treats the
fluid as a mixture of perfect gases; the corresponding conservation
equations for an open domain are
\begin{eqnarray*}
\frac{\partial (\rho \boldsymbol{u})}{\partial t} + 
\nabla \cdot \left(\rho  \boldsymbol{u} \boldsymbol{u} + \tau \right)
&=& -\nabla \pi + \rho \, \boldsymbol{F}  ,
\label{eq:mom}
\\
\frac{\partial (\rho Y_i)}{\partial t} +
\nabla \cdot \left( \rho Y_i \boldsymbol{u} + \boldsymbol{\mathcal{F}}_{i} \right)
&=& \rho \, \dot{\omega}_i,
\label{eq:species}
\\
\frac{ \partial (\rho h)}{ \partial t} +
\nabla \cdot \left( \rho h \boldsymbol{u} + \boldsymbol{\mathcal{Q}} \right) &=& 0 ,
\label{eq:enthalpy}
\end{eqnarray*}
where $\rho$ is the density, $\boldsymbol{u}$ is the velocity, $h$ is the mass-weighted enthalpy,
$T$ is temperature.  $Y_i$ is the mass fraction and $\dot{\omega}_i$ is the molar
production rate for species $i$.
A mixture-averaged model is assumed for diffusive transport, ignoring
Dufour and Soret effects; $\tau$ is the stress tensor; 
$\boldsymbol{\mathcal{Q}}$$\,=\,$$h_{i} \boldsymbol{\mathcal{F}}_{i} - \lambda \nabla T $ 
is the heat flux and 
$\boldsymbol{\mathcal{F}}_i$$\,=\,$$- \rho \boldsymbol{\mathcal{D}}_i \nabla X_{i}$ 
is the species diffusion.
The momentum source term, $\boldsymbol{F}$, is a long-wavelength
forcing term designed to establish and maintain turbulence with the desired properties.

The chemical kinetics and transport are modelled 
using the dodecane mechanism of You~\etal~\cite{YouWang2009}, 
consisting of 56 species with 289 fundamental reactions.
These evolution equations are supplemented by
%an equation of state for a perfect gas mixture, and
CHEMKIN-compatible databases for thermodynamic quantities.
%species enthalpy, $h_i(T)$, and
%transport coefficients, 
%$\lambda$, and $\mathcal{D}_i$ (specified as part of the kinetics model).
Transport properties are computed using EGLIB \cite{EGLIB}.

The basic discretisation combines a spectral deferred corrections (SDC) coupling
of chemistry and transport \cite{Nonaka2012} with a density-weighted approximate
projection method for low Mach number flow \cite{DayBell2000}.  The projection method
implements a constrained evolution on the velocity field via the SDC iterations,
which ensures that the update simultaneously satisfies the 
equation of state and discrete conservation of mass and total enthalpy.
A time-explicit approach is used for advection; faster diffusion
and chemistry processes are treated time-implicitly, and iteratively
coupled together within the defferred corrections strategy.
Since the low Mach system does not support acoustic waves,
the time step size is governed by a CFL constraint based on advective transport.
The integration algorithm is second-order accurate in space
and time.

The performance of the numerical scheme for direct numerical 
simulation of premixed flame systems in regimes comparable to the present study
was examined in \cite{Aspden08b}.  An {\em effective} Kolmogorov
length scale, $\eta_{\mbox{\it eff}}$, was formulated, which measures the actual Kolmogorov length
scale realised in a simulation at a given resolution.
Here, the most computationally demanding simulation, having the highest turbulence levels,
has a computational cell width that is approximately 1.27 times the
Kolmogorov length scale, $\eta$.  In this case, the numerical scheme produces
$\eta_{\mbox{\it eff}}/\eta\,$$<\,$1.03.  All other cases were better resolved.
